\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{hyperref}

\usepackage{listings}
\usepackage{rotating, graphicx}
\graphicspath{{./}, {./image/}}
\usepackage{booktabs, natbib}
% \usepackage{breakurl}
% \usepackage [english]{babel}
\usepackage{amsmath, amsbsy, amsthm, epsfig, epsf, psfrag, graphicx,
amssymb, enumerate}
\usepackage{bm}
\usepackage{multirow, multicol}

\usepackage[dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.1, 0.2, 0.6}


\newcommand{\jy}[1]{\textcolor{orange}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\of}[1]{\textcolor{purple}{(OF: #1)}}

\sloppy

% \usepackage{csquotes}
% \usepackage [autostyle, english = american]{csquotes}
% \MakeOuterQuote{"}

% \usepackage{bibentry}
\newenvironment{comment}%
{\begin{quotation}\noindent\small\it\color{darkblue}\ignorespaces%
}{\end{quotation}}


\begin{document}

\begin{center}
  {\Large\bf TAS-23-296.R1: Response to the Comments}
\end{center}

We sincerely thank the Editor for tentatively accepting our submission
and Reviewer~1 for the additional suggestions. We are pleased to
present our final revisions.


The manuscript has been revised as follows, with the most notable
changes being:
\begin{enumerate}
\item
  Restructuring: The manuscript is now organized into four
  sections: Introduction, Analysis of 2022 Times, Analysis of a Proper
  RT Barrier, and Discussion. The second and third sections
  further include subsections for Data, Methods, and Results to
  enhance clarity and readability.
\item
  General Refinements: The manuscript has been carefully revised to
  improve clarity, coherence, and overall presentation.
\end{enumerate}

Point-by-point responses to the comments are as follows, with the
original comments quoted in \emph{\color{darkblue} italic blue}.


\subsection*{To Reviewer 1}

\begin{comment}
This revision is a solid improvement on the original paper exploring the
anomalous reaction times at the 2022 World Championships and the implications of
using 0.100 seconds as the cut off for acceptable reaction times. The updated
paper addresses many of my original points of feedback, including, notably, data
from more events, improving the statistical power and generalizability of the
results. I also think the expansion of the features considered in the model is
beneficial (although, I do think considering a Bayesian approach to the
distribution estimation could be beneficial).

However, I still believe there is room to improve the overall quality of the
paper, both from a technical/methodological perspective and from a more general
clarity and organization perspective.


\end{comment}

Thank you for the thoughtful feedback and for recognizing the
improvements in our revision. The remaining points are well taken, and
we address them below.

\begin{comment}
First, I think the description of the data would benefit from an easier to
follow organization system. Currently, headings repeat and the overall
difference between the usage of data from the World Championships in one setting
instead of the other is not as easily understood as it could or probably should
be. I would recommend a more intuitive organization of this material and, if
possible, the construction and inclusion of a table summarizing the various data
details in each analysis. As a reader, I would like a fast way to keep track of
the important details in each case.
\end{comment}

Distinguishing between the datasets used in each section was a valid
concern. To improve clarity, the paper has been restructured so that
each problem’s data, methods, and results appear consecutively, making
it easier to follow. The dataset for the 2022 analysis is now
described in the second paragraph of Section~2.1.1, while the dataset
for the RT barrier analysis is introduced in the first paragraph of
Section~3.1.


\begin{comment}
Second, I think the results themselves would benefit from organizing around the
question or problem being addressed instead of the method. Organizing by method
makes it harder for your reader to connect the results to the research questions
or prompts.
\end{comment}

This was an excellent suggestion, and we appreciate the insight. To
improve clarity and better align the results with the research
questions, the paper has been reorganized. Section~2 now focuses
on whether athletes competed faster in 2022, while Section~3
examines the RT barrier. The Introduction and Discussion
sections provide context and highlight the connection between these
two analyses.


\begin{comment}
Third, I think the manuscript as a whole needs some general editing and cleaning
up before publication. For example, there are discrepancies in the language used
in some of the tables - in Table 1 the groups are labeled both "Men/Women" and
"Men/Female" when consistency would be helpful for the readers. Similarly, Table
2 uses the model notation to as column headers, making it hard to quickly assess
what it is communicating. A stronger organizational approach could also help the
overall clarity of the article.
\end{comment}

The manuscript has been further refined for clarity and
consistency. The language in Table~1 is now uniform. With the
restructuring, the definitions of the symbols in Table~2 are now
included within the same section, making them easier to connect.

\begin{comment}
Finally, I think it is important that the article consider the possible causes
of differences between the competitions that are not accounted for in the
matched-pairs design. As I understand the approach, the analysis does not
account for two important potential sources of change in the reaction times::
seasonality and age. Seasonality could come into play if athletes' reaction
times improve as they attempt to peak in their season (likely at the highest
level competition they will participate in). If athletes expect to compete at
the World Championships, they might not be as fit or prepared at their national
championships and produce slower reaction times. Similarly, age could lead to
differences in an athletes' reaction times as older athletes will react slower
than younger athletes. While it would be great to see these features accounted
for directly, I think at a minimum they and their likely influences on the
results, should be discussed in the overall limitations of the analysis.
\end{comment}

We appreciate this thoughtful comment and have addressed it in the
discussion section. The revised text clarifies that comparisons across
multiple years (2019, 2022, and 2023) help mitigate the effects of
seasonality and age, as athletes contribute to different pairings
across years. Additionally, while seasonal peaking may influence RTs
at national competitions, the inclusion of both national and
international events provides a broader competitive context, reducing
the likelihood that seasonality alone explains the observed
differences. This clarification has been added to the first paragraph
of the discussion.

\begin{comment}
Altogether, I think the technical work here is solid and shows that the reaction
times at the 2022 World Championships were identifiably faster than at other
major competitions and that 0.100 threshold for acceptable reaction times is
likely too conservative of a standard given athletes' natural abilities. The
largest areas of need are mainly in the organization and overall clarity of the
paper.
\end{comment}

We are delighted that the technical work
is recognized as solid. The revisions have focused on improving the
organization and clarity of the paper to better communicate the key
findings.


% \bibliographystyle{apalike}
% \bibliography{citations}
\end{document}
