---
title: "GLMM"
author: "Blinded"
date: "2023-12-01"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### This is the code file that shows our results. Code chunks have been labeled for convience.  

All outputs rely on code chunks 1-5, run these first.
To see the outputs of Table 2 run code chunks 6-8
To see the outputs of Table 2 run code chunks 11-14
To see the outputs of Table 4 run code chunks 15-17


### 1
The code below loads in the data sets from the csv file.
```{r}
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")

with2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles")

without2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022)
```


### 2
This code below creates the null models without any random effects.  These are
used as a baseline to compare the venue effect models to in the simulation code
below.
```{r}
library(lme4)
without2022_one_gamma_log <- glm(RxnTime ~ 1, data = without2022, 
                                    family = Gamma(link = "log"))
with2022_one_gamma_log <- glm(RxnTime ~ 1, data = with2022, 
                                family = Gamma(link = "log"))
```

### 3
This code creates the venue effect only models.
```{r}
without2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = without2022, 
                                       family = Gamma(link = "log"))
with2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = with2022, 
                                   family = Gamma(link = "log"))
```


### 4
This creates the heat effect only models.
```{r}
without2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = without2022, 
                             family = Gamma(link = "log"))
with2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = with2022, 
                         family = Gamma(link = "log"))
```


### 5
This creates the models with a venue effect and heat effect.
```{r}
without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                                   data = without2022,
                                   family = Gamma(link = "log"))
with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = with2022,
                               family = Gamma(link = "log"))
```


### 6
This shows a summary of the venue effect only models.
```{r}
summary(with2022_venue_gamma_log)
summary(without2022_venue_gamma_log)
```


### 7
This shows a summary of the heat effect only models.
```{r}
summary(with2022_heat_gamma_log)
summary(without2022_heat_gamma_log)
```


### 8
This shows a summary of the two effect models.
```{r}
summary(with2022_two_effect_gamma_log)
summary(without2022_two_effect_gamma_log)
```


### 9
This test compares the full model with the heat effect only model.
```{r}
anova(with2022_heat_gamma_log, with2022_two_effect_gamma_log)
```
This anova test determined that the two effect model is an improvement over the
heat effect model. The p-value is $1.21 \cdot 10^{-5}$ which is significant at 
$\alpha = 0.05$.


### 10
This test compares the full model with the venue effect only model.
```{r}
anova(with2022_venue_gamma_log, with2022_two_effect_gamma_log)
```
This anova test determined that the two effect model is a significant
improvement over the Venue effect model.

We ran additional tests and found that including `Event` as a predictor in the
model did not significantly improve the model.


### 11
This code simulates the probability of observing a reaction time less than 0.1
for the various data sets.  The function takes in two models and a reaction
time, it is important that the two models that are compared are from the same
data set; we want to compare `old_finals_two_effect` and 
`old_finals_one_gamma_log`.
```{r}
sim_two_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]
    
    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)
    
    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)
    
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }
  
  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
}
```


### 12
This next chunk of code runs the simulation code for each dataset comparing the
two effect model to the base model. We run this code at 3 reaction time barriers:
0.08, 0.09, and 0.1 seconds.
```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.1)
sim_two_effect_time(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.1)
```


### 13
```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.09)
sim_two_effect_time(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.09)
```


### 14
```{r}
set.seed(1)
sim_two_effect_time(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.08)
sim_two_effect_time(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.08)
```


### 15
`sim_two_effect_prob` is a slight modification of the earlier function that
instead takes as input a probability rather than a time.  The output is a
reaction time barrier that allows a reaction time below the barrier one in every
1/(input probability) times. Decreasing the probability will impose a stricter
barrier as it is less probable to observe a reaction time of 0.08 rather than
0.09.  Below are the recommended reaction time barriers to disqualify on average
1 in every 1,000 starts.
```{r}
sim_two_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]
    
    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)
    
    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)
    
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }
  
  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```


### 16
This runs the `sim_two_effect_prob` function at a probability of 1 in every 
1,000 sprinters.
```{r}
#9
set.seed(1)
sim_two_effect_prob(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.001)
sim_two_effect_prob(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.001)
```


### 17
The code below repeats the previous section but for an average of 1 in every
10,000 sprinters.
```{r}
set.seed(1)
sim_two_effect_prob(without2022_two_effect_gamma_log,
                    without2022_one_gamma_log, 0.0001)
sim_two_effect_prob(with2022_two_effect_gamma_log,
                    with2022_one_gamma_log, 0.0001)
```


### 18
We now examine the men's 100 meter dash times from 1999 to 2023.
```{r}
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
dash_2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
                   & Event == "Dash")
dash_no_2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
                   & Event == "Dash" & Venue != 2022)
```

```{r}
library(tidyverse)
dash_2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

dash_no_2022 <- times |>
  filter(Venue != 2022) |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```



### 19
Baseline models
```{r}
library(lme4)
dash_no_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash_no_2022, 
                                    family = Gamma(link = "log"))
dash_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash_2022, 
                                family = Gamma(link = "log"))
```


### 20
This code creates the venue effect only models.
```{r}
dash_no_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_no_2022, 
                                       family = Gamma(link = "log"))
dash_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_2022, 
                                   family = Gamma(link = "log"))
```


### 21
This creates the heat effect only models.
```{r}
dash_no_2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = dash_no_2022, 
                             family = Gamma(link = "log"))
dash_2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = dash_2022, 
                         family = Gamma(link = "log"))
```


### 22
This creates the models with a venue effect and heat effect.
```{r}
dash_no_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                                   data = dash_no_2022,
                                   family = Gamma(link = "log"))
dash_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = dash_2022,
                               family = Gamma(link = "log"))
```


### 23
Now we test
```{r}
sim_two_effect_time(dash_no_2022_two_effect_gamma_log,
                    dash_no_2022_one_gamma_log, 0.1)
sim_two_effect_time(dash_2022_two_effect_gamma_log,
                    dash_2022_one_gamma_log, 0.1)
```
## 24
In response to reviewer 2's comments that questioned the probability of observing
the venue effect for 2022 we perform the following analysis. We store the venue
effect for 2022 in `effect2022` and we store the standard deviation in `sd2022`
```{r}
library(tidyverse)
ranef_results <- data.frame(ranef(with2022_two_effect_gamma_log))
venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
venues$year <- as.character(venues$grp)
venues$year <- as.numeric(venues$year)
effect2022 <- venues |> filter(grp == 2022) |> select(condval) |> pull()
sd2022 <- as.data.frame(VarCorr(with2022_two_effect_gamma_log)) |> filter(grp == "Venue") |> select(sdcor)|> pull()

```
Now we can calculate how extreme this value is
```{r}
pnorm(effect2022, mean = 0, sd = sd2022)
```

## Now we look at the 200 meter dash results
```{r}
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

```{r}
library(tidyverse)
dash200_2022 <- times |>
  filter(Event == "200 Dash") |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

dash200_no_2022 <- times |>
  filter(Event == "200 Dash") |>
  filter(Venue != 2022) |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

```{r}
library(lme4)
dash200_no_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash200_no_2022, 
                                    family = Gamma(link = "log"))
dash200_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash200_2022, 
                                family = Gamma(link = "log"))

dash200_no_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                                   data = dash200_no_2022,
                                   family = Gamma(link = "log"))
dash200_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = dash200_2022,
                               family = Gamma(link = "log"))
```
Now we test
```{r}
sim_two_effect_time(dash200_no_2022_two_effect_gamma_log,
                    dash_no_2022_one_gamma_log, 0.1)
sim_two_effect_time(dash200_2022_two_effect_gamma_log,
                    dash_2022_one_gamma_log, 0.1)
```
These are low probabilities indicating that a reaction time below 0.1 seconds
is unlikely for the 200 meter dash.
```{r}
library(ggplot2)
ggplot(dash200_2022, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.25),
                            breaks = seq(0.1, 0.25, 0.05)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
```



## We will compare our results to the womens data starting with hurdles and then dash

```{r}
library(tidyverse)
women_hurdles <- times |>
  filter(Event == "100 Hurdles") |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "F") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

women_dash <- times |>
  filter(Event == "100 Dash") |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "F") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

```{r}
library(ggplot2)
ggplot(women_hurdles, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.25),
                            breaks = seq(0.1, 0.25, 0.05)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(2000,2024),
                                              breaks = seq(2001, 2024, 2))
```
```{r}
library(ggplot2)
ggplot(women_dash, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.25),
                            breaks = seq(0.1, 0.25, 0.05)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
```



```{r}
men_dash <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
```

